Beyond Ultra-Diffuse Galaxies I: Mass-Size Outliers Among the Satellites of Milky Way Analogs

### 2022-10-25 paper1.tex
fixed numbers of contam

\begin{deluxetable*}{lcc}
\tablecaption{LSBG search steps and the number of remaining objects after each step}
\tablewidth{20cm}
\label{tab:steps_flow}
\tablehead{
\colhead{Process} & \colhead{Description} &
\colhead{Remaining objects}
}
\startdata
Initial detection & \S\ref{sec:detection} & 86,002 \\
Matching with MW analogs at $0.01 < z < 0.04$ & \S\ref{sec:match} & 10,579 \\
Deblending &\S\ref{sec:deblending} & 2,673\\
Spergel profile modeling & \S\ref{sec:modeling}, \S\ref{sec:match} & 2,510\\
Completeness & \S\ref{sec:comp_meas},\S\ref{ap:comp_meas_unc} & 2,510 \\
Measurement error and uncertainty & \S\ref{sec:comp_meas},\S\ref{ap:comp_meas_unc} & 2,510\\
Mass-size outlier selection & \S\ref{sec:sample} & 362 (UPG), 432 (UDG)
\enddata
\end{deluxetable*}

@@

% However, a significant fraction of spurious objects remain in this cross-matched LSBG sample, including galactic cirrus, tidal streams, shredded large galaxies, and compact sources blended with galaxy outskirts. 
As described in \S\ref{sec:deblending}, we then perform a deblending step to effectively remove spurious objects. There are 2,673 objects left after the deblending cuts. Next, we model these remaining objects using the Spergel profile as described in \S\ref{sec:modeling} and obtain their photometry and structural parameters from the best-fit models. We also remove duplicate objects. In the final catalog, we are left with 2,510 LSBG candidates associated with 689 MW analogs as our final LSBG sample. The host number drops to 689 because many hosts in the prior sample has no true positives as their satellites. We note that these LSBGs are matched to MW hosts in projection but do not have direct distance measurements, and we have not applied any statistical background correction at this stage.

@@

We randomly select a continuous patch of sky of 24 deg$^{2}$ in HSC PDR2 regardless of whether it contains MW analogs\footnote{$345\ \deg < \rm{R.A.} < 351\ \deg$, $-1.5\ \deg < \rm{Dec} < 2.5\ \deg$, which is not dominated by Galactic cirrus and stars. Such a patch of sky with 24 deg$^2$ is already enough for us to derive a background contamination fraction with small Poisson noise.}. Then we perform the same deblending and modeling steps for the 2,707 LSBG candidates detected in this region. We also remove objects that are already in our mass-size outlier samples and objects with completeness less than 0.1. We also did the same visual inspection as for UDGs and UPGs to remove false positives. In the end, there are 480 LSBGs (excluding false positives) representing a population of possible contaminants for the UDG and UPG samples. Next, because both UDG and UPG are defined based on a physical size, we randomly associate the 480 LSBGs with the hosts of our UDGs and UPGs, respectively. Then we calculate a physical size for each LSBG and classify an LSBG as a ``fake`` UDG or UPG if it satisfies the corresponding definition. We repeat such random matching 200 times. In the end, we obtain 11,458 fake UDGs and 11,867 fake UPGs. These UDGs and UPGs are ``fake'' only in the sense of being associated with random MW hosts.

In this way, the number density of fake UDGs is estimated to be $S_{\rm UDG} = 1.94\pm0.03\ \mathrm{deg}^{-2}$, and the number density of fake UPGs is $S_{\rm UPG} = 2.12\pm0.03\ \mathrm{deg}^{-2}$. The catalogs of fake UDG and UPGs are also available online\footnote{\url{https://astrojacobli.github.io/research/BeyondUDG/}}. Using the number densities and the probed area of our survey, we calculate the contamination fraction for both the UDG and UPG samples. For the UDG sample, we find $f_{\rm contam}^{\rm UDG} \approx 16\%$; for the UPG sample, the contamination fraction is $f_{\rm contam}^{\rm UPG} \approx 20\%$. We take the contamination fraction into account when calculating the abundances in \S\ref{sec:results}.

@@

We define the UDG (UPG) abundance as the average number of UDGs (UPGs) per host galaxy. We searched 689 MW analogs. In our UDG sample, there are 412 UDGs associated with 258 hosts. After correcting for background contamination (see \S \ref{sec:bkg}) and completeness (see \S \ref{sec:comp_meas}), the UDG abundance of MW analogs is $N_{\rm UDG} = 0.76\pm 0.04$ per host. The UPG abundance in our sample is $N_{\rm UPG} = 0.61\pm 0.04$ per host. Here we neglect the fact that the number of satellites contained within the virial sphere is different from the number of satellite within a projected cylinder with virial radius (the so-called deprojection factor, \citealt{vdBurg2017}).

@@

Another interesting quantity is the fraction of satellites that are mass-size outliers in a group or cluster. This fraction represents how efficient an environment is at producing puffy satellites. We denote this quantity as the UDG (UPG) fraction hereafter, which characterizes the tail of the size distribution of satellites. In order to calculate this fraction, we assume that MW analogs have $\sim 6 \pm 1$ satellites with $M_\star > 10^{6.5}\ M_\odot$ \citep{CarlstenELVES2022}. This lower limit in stellar mass is roughly our detection limit (see Fig. \ref{fig:mass_size}). Then for our samples, the UDG (UPG) fraction is $f_{\rm UDG} \approx 0.12 \pm 0.03$ and $f_{\rm UPG} \approx 0.10 \pm 0.03$. In ELVES, among 351 satellites with secure distances in 30 MW analogs, we identify 13 UDGs associated with 10 hosts, and 36 UPGs associated with 20 hosts. Thus the UDG fraction in ELVES is $f_{\rm UDG}^{\rm ELVES} \approx 0.04 \pm 0.01$, the UPG fraction in ELVES is $f_{\rm UPG}^{\rm ELVES} \approx 0.10 \pm 0.02$. We further select ELVES galaxies with $M_\star > 10^{6.5}\ M_\odot$ and find the same UDG and UPG fractions. We also compare our results with results in the Virgo cluster. Taking the data from the Next Generation Virgo Cluster Survey (NGVS, \citealt{Ferrarese2020}) and only considering satellites with $M_\star > 10^{6.5}\ M_\odot$, there are 14 UDGs and 17 UPGs out of 218 identified satellites (with no completeness correction). Thus their UDG (UPG) fraction is $f_{\rm UDG}^{\rm Virgo} \approx 0.06\pm 0.02$ and $f_{\rm UPG}^{\rm Virgo} \approx 0.08\pm 0.02$. However, we note that \citet{Ferrarese2020} only survey the central 4 deg$^2$ of the Virgo cluster, which probably biases the UDG (UPG) fraction. It seems that our UDG abundance and fraction are higher than both ELVES and NGVS. Our UPG abundance and fraction are consistent with ELVES and NVGS. In the future, a systematic and careful comparison on UDG (UPG) fraction in different environments is needed.

>>>>

\begin{deluxetable*}{lcc}
\tablecaption{LSBG search steps and the number of remaining objects after each step}
\tablewidth{20cm}
\label{tab:steps_flow}
\tablehead{
\colhead{Process} & \colhead{Description} &
\colhead{Remaining objects}
}
\startdata
Initial detection & \S\ref{sec:detection} & 86,002 \\
Matching with MW analogs at $0.01 < z < 0.04$ & \S\ref{sec:match} & 10,579 \\
Deblending &\S\ref{sec:deblending} & 2,673\\
Spergel profile modeling & \S\ref{sec:modeling}, \S\ref{sec:match} & 2,510\\
Completeness & \S\ref{sec:comp_meas},\S\ref{ap:comp_meas_unc} & 2,510 \\
Measurement error and uncertainty & \S\ref{sec:comp_meas},\S\ref{ap:comp_meas_unc} & 2,510\\
Mass-size outlier selection & \S\ref{sec:sample} & 337 (UPG), 412 (UDG)
\enddata
\end{deluxetable*}

@@

% However, a significant fraction of spurious objects remain in this cross-matched LSBG sample, including galactic cirrus, tidal streams, shredded large galaxies, and compact sources blended with galaxy outskirts. 
As described in \S\ref{sec:deblending}, we then perform a deblending step to effectively remove spurious objects. There are 2,673 objects left after the deblending cuts. Next, we model these remaining objects using the Spergel profile as described in \S\ref{sec:modeling} and obtain their photometry and structural parameters from the best-fit models. We also remove duplicate objects. In the final catalog, we are left with 2,510 LSBG candidates associated with 689 MW analogs as our final LSBG sample. These MW analogs occupy a sky area of 75.94 deg$^{2}$ (out to $1\ R_{\rm vir}$). The host number drops to 689 because many hosts in the prior sample has no true positives as their satellites. We note that these LSBGs are matched to MW hosts in projection but do not have direct distance measurements, and we have not applied any statistical background correction at this stage.

@@

We randomly select a continuous patch of sky of 24 deg$^{2}$ in HSC PDR2 regardless of whether it contains MW analogs\footnote{$345\ \deg < \rm{R.A.} < 351\ \deg$, $-1.5\ \deg < \rm{Dec} < 2.5\ \deg$, which is not dominated by Galactic cirrus and stars. Such a patch of sky with 24 deg$^2$ is already enough for us to derive a background contamination fraction with small Poisson noise.}. Then we perform the same deblending and modeling steps for the 2,707 LSBG candidates detected in this region. We also remove objects that are already in our mass-size outlier samples and objects with completeness less than 0.1. We also did the same visual inspection as for UDGs and UPGs to remove false positives. In the end, there are 480 LSBGs (excluding false positives) representing a population of possible contaminants for the UDG and UPG samples. Next, because both UDG and UPG are defined based on a physical size, we randomly associate the 480 LSBGs with the 689 MW analogs that are surveyed in this paper. Then we calculate a physical size for each LSBG and classify an LSBG as a ``fake`` UDG or UPG if it satisfies the corresponding definition. We repeat such random matching 200 times. In the end, we obtain 8,392 fake UDGs and 8,485 fake UPGs. These UDGs and UPGs are ``fake'' only in the sense of being associated with random MW hosts.

In this way, the number density of fake UDGs is estimated to be $S_{\rm UDG} = 1.75\pm0.03\ \mathrm{deg}^{-2}$, and the number density of fake UPGs is $S_{\rm UPG} = 1.70\pm0.03\ \mathrm{deg}^{-2}$. The catalogs of fake UDG and UPGs are also available online\footnote{\url{https://astrojacobli.github.io/research/BeyondUDG/}}. Using the number densities and the probed area of our survey, we calculate the contamination fraction for both the UDG and UPG samples. For the UDG sample, we find $f_{\rm contam}^{\rm UDG} \approx 32\%$; for the UPG sample, the contamination fraction is $f_{\rm contam}^{\rm UPG} \approx 38\%$. We take the contamination fraction into account when calculating the abundances in \S\ref{sec:results}.

% For the UDG sample, we find $f_{\rm contam}^{\rm UDG} \approx 36\%$; for the UPG sample, the contamination fraction is $f_{\rm contam}^{\rm UPG} \approx 47\%$. We take the contamination fraction into account when calculating the abundances in \S\ref{sec:results}.

@@

We define the UDG (UPG) abundance as the average number of UDGs (UPGs) per host galaxy. We searched 689 MW analogs. In our UDG sample, there are 412 UDGs associated with 258 hosts. After correcting for background contamination (see \S \ref{sec:bkg}) and completeness (see \S \ref{sec:comp_meas}), the UDG abundance of MW analogs is $N_{\rm UDG} = 0.61\pm 0.04$ per host. The UPG abundance in our sample is $N_{\rm UPG} = 0.47\pm 0.04$ per host. Here we neglect the fact that the number of satellites contained within the virial sphere is different from the number of satellite within a projected cylinder with virial radius (the so-called deprojection factor, \citealt{vdBurg2017}).

@@

Another interesting quantity is the fraction of satellites that are mass-size outliers in a group or cluster. This fraction represents how efficient an environment is at producing puffy satellites. We denote this quantity as the UDG (UPG) fraction hereafter, which characterizes the tail of the size distribution of satellites. In order to calculate this fraction, we assume that MW analogs have $\sim 6 \pm 1$ satellites with $M_\star > 10^{6.5}\ M_\odot$ \citep{CarlstenELVES2022}. This lower limit in stellar mass is roughly our detection limit (see Fig. \ref{fig:mass_size}). Then for our samples, the UDG (UPG) fraction is $f_{\rm UDG} \approx 0.10 \pm 0.03$ and $f_{\rm UPG} \approx 0.08 \pm 0.03$. In ELVES, among 351 satellites with secure distances in 30 MW analogs, we identify 13 UDGs associated with 10 hosts, and 36 UPGs associated with 20 hosts. Thus the UDG fraction in ELVES is $f_{\rm UDG}^{\rm ELVES} \approx 0.04 \pm 0.01$, the UPG fraction in ELVES is $f_{\rm UPG}^{\rm ELVES} \approx 0.10 \pm 0.02$. We further select ELVES galaxies with $M_\star > 10^{6.5}\ M_\odot$ and find the same UDG and UPG fractions. We also compare our results with results in the Virgo cluster. Taking the data from the Next Generation Virgo Cluster Survey (NGVS, \citealt{Ferrarese2020}) and only considering satellites with $M_\star > 10^{6.5}\ M_\odot$, there are 14 UDGs and 17 UPGs out of 218 identified satellites (with no completeness correction). Thus their UDG (UPG) fraction is $f_{\rm UDG}^{\rm Virgo} \approx 0.06\pm 0.02$ and $f_{\rm UPG}^{\rm Virgo} \approx 0.08\pm 0.02$. However, we note that \citet{Ferrarese2020} only survey the central 4 deg$^2$ of the Virgo cluster, which probably biases the UDG (UPG) fraction. It seems that our UDG abundance and fraction are higher than both ELVES and NGVS. Our UPG abundance and fraction are consistent with ELVES and NVGS. In the future, a systematic and careful comparison on UDG (UPG) fraction in different environments is needed.
